Equations & R Code

### 2016-10-15 manuscript/coexistence-stability_ms.Rmd
updates ms intro with research questions

INTRODUCTION
------------
@MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized that species can compensate for each other and stabilize functioning in ecosystems subject to temporal variation in environmental conditions. 
This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests stability increases with diversity because species respond dissimilarly to environmental conditions – species A has highest growth rates under conditions X whereas species B has highest growth rates under conditions Y. 
More species confer temporal stability by broadening the range of conditions under which the community maintains function [@Loreau2010].
Diverse models all predict a positive relationship between species richness and ecosystem stability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].
However, empirical support for a positive diversity-stability relationship is not unequivocal [@Jiang2009].
Furthermore, theory on the relationship between biodiversity and ecosystem stability often ignores the processes that determine species coexistence in variable environments [Loreau 2010, but see @Chesson2001].

@@

To reconcile these two bodies of theory, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
We focus on the storage effect and relative nonlinearity using a general consumer-resource model.
First, we use model simulations to investigate the diversity-stability relationship across a gradient of environmental variability.
Counter to common expectations, we find that a negative diversity-stability relationship should be expected when sampling occurs over natural diversity gradients and species coexistence is fluctuation dependent.
Importantly, and in line with previous theory [@Ives2002b; @Loreau2010; @DeMazancourt2013], at a given level of environmental variability, increasing species richness stabilizes ecosystem function, even when coexistence is fluctuation dependent.

Second, we explore the net effect of increasing environmental variability on ecosystem stability by isolating the gain in stability due to increased richness and the loss in stability due to increased amplitude of species fluctuations.
We find that increasing environmental variability does not always lead to a decrease in ecosystem stability: environmental variability promotes species richness, which stabilizes ecosystem functioning via portfolio effects even as the environment becaomes more variable.
Whether or not increasing environmental variability results in species gains depends on the specific coexistence mechanism, the traits of resident species, and the traits of the species in the regional pool.

>>>>

INTRODUCTION
------------
@MacArthur1955, @Elton1958, and even Darwin [@Turnbull2013] recognized that species can compensate for each other and stabilize functioning in ecosystems subject to temporal variation in environmental conditions. 
This idea underlies the "insurance hypothesis" [@Yachi1999], which suggests stability increases with diversity because species respond dissimilarly to environmental conditions – species A has highest growth rates under conditions X whereas species B has highest growth rates under conditions Y. 
More species confer temporal stability by broadening the range of conditions under which the community maintains function [@Loreau2010].
Diverse models all predict a positive relationship between species richness and ecosystem stability [@Lehman2000; @Ives2002b; @Loreau2013], and experimental tests tend to support such a prediction [@Tilman2006; @Hector2010].

However, the ability of biodiversity-ecosystem functioning (BEF) experiments to represent real-world dynamics is the subject of debate [@Wardle2016; @Eisenhauer2016].
Much of the debate centers around the fact that species losses in BEF experiments are not allowed to be offset by species gains.
Theory typically suffers from the same limitation, in that it often ignores the processes that could lead to species gains in temporally variable environments [Loreau 2010, but see @Chesson2001].

@@

To reconcile these two bodies of theory, we extend theory on the relationship between species richness and ecosystem stability to cases in which species coexistence explicitly depends on environmental fluctuations and species-specific responses to environmental conditions.
We focus on the storage effect and relative nonlinearity using a general consumer-resource model.
We use the model to investigate two questions:

1. Does the direction of the diversity-stability relationship remain positive when species coexistence is fluctuation-dependent? And,

2. When species coexistence is fluctuation-dependent, how does increasing environmental variability impact ecosystem stability?

### 2016-10-15 manuscript/coexistence-stability_ms.tex
updates ms intro with research questions

MacArthur (1955), Elton (1958), and even Darwin (Turnbull et al. 2013)
recognized that species can compensate for each other and stabilize
functioning in ecosystems subject to temporal variation in environmental
conditions. This idea underlies the ``insurance hypothesis'' (Yachi and
Loreau 1999), which suggests stability increases with diversity because
species respond dissimilarly to environmental conditions -- species A
has highest growth rates under conditions X whereas species B has
highest growth rates under conditions Y. More species confer temporal
stability by broadening the range of conditions under which the
community maintains function (Loreau 2010). Diverse models all predict a
positive relationship between species richness and ecosystem stability
(Lehman and Tilman 2000, Ives and Hughes 2002, Loreau and {{de
Mazancourt}} 2013), and experimental tests tend to support such a
prediction (Tilman et al. 2006, Hector et al. 2010). However, empirical
support for a positive diversity-stability relationship is not
unequivocal (Jiang and Pu 2009). Furthermore, theory on the relationship
between biodiversity and ecosystem stability often ignores the processes
that determine species coexistence in variable environments (Loreau
2010, but see Chesson et al. 2001).

@@

To reconcile these two bodies of theory, we extend theory on the
relationship between species richness and ecosystem stability to cases
in which species coexistence explicitly depends on environmental
fluctuations and species-specific responses to environmental conditions.
We focus on the storage effect and relative nonlinearity using a general
consumer-resource model. First, we use model simulations to investigate
the diversity-stability relationship across a gradient of environmental
variability. Counter to common expectations, we find that a negative
diversity-stability relationship should be expected when sampling occurs
over natural diversity gradients and species coexistence is fluctuation
dependent. Importantly, and in line with previous theory (Ives and
Hughes 2002, Loreau 2010, Mazancourt et al. 2013), at a given level of
environmental variability, increasing species richness stabilizes
ecosystem function, even when coexistence is fluctuation dependent.

Second, we explore the net effect of increasing environmental
variability on ecosystem stability by isolating the gain in stability
due to increased richness and the loss in stability due to increased
amplitude of species fluctuations. We find that increasing environmental
variability does not always lead to a decrease in ecosystem stability:
environmental variability promotes species richness, which stabilizes
ecosystem functioning via portfolio effects even as the environment
becaomes more variable. Whether or not increasing environmental
variability results in species gains depends on the specific coexistence
mechanism, the traits of resident species, and the traits of the species
in the regional pool.

@@

@@

Mazancourt, C. de, F. Isbell, A. Larocque, F. Berendse, E. {De Luca}, J.
B. Grace, B. Haegeman, H. {Wayne Polley}, C. Roscher, B. Schmid, D.
Tilman, J. van Ruijven, A. Weigelt, B. J. Wilsey, and M. Loreau. 2013.
Predicting ecosystem stability from community composition and
biodiversity. Ecology Letters 16:617--625.

@@

>>>>

MacArthur (1955), Elton (1958), and even Darwin (Turnbull et al. 2013)
recognized that species can compensate for each other and stabilize
functioning in ecosystems subject to temporal variation in environmental
conditions. This idea underlies the ``insurance hypothesis'' (Yachi and
Loreau 1999), which suggests stability increases with diversity because
species respond dissimilarly to environmental conditions -- species A
has highest growth rates under conditions X whereas species B has
highest growth rates under conditions Y. More species confer temporal
stability by broadening the range of conditions under which the
community maintains function (Loreau 2010). Diverse models all predict a
positive relationship between species richness and ecosystem stability
(Lehman and Tilman 2000, Ives and Hughes 2002, Loreau and {{de
Mazancourt}} 2013), and experimental tests tend to support such a
prediction (Tilman et al. 2006, Hector et al. 2010).

However, the ability of biodiversity-ecosystem functioning (BEF)
experiments to represent real-world dynamics is the subject of debate
(Wardle 2016, Eisenhauer et al. 2016). Much of the debate centers around
the fact that species losses in BEF experiments are not allowed to be
offset by species gains. Theory typically suffers from the same
limitation, in that it often ignores the processes that could lead to
species gains in temporally variable environments (Loreau 2010, but see
Chesson et al. 2001).

@@

To reconcile these two bodies of theory, we extend theory on the
relationship between species richness and ecosystem stability to cases
in which species coexistence explicitly depends on environmental
fluctuations and species-specific responses to environmental conditions.
We focus on the storage effect and relative nonlinearity using a general
consumer-resource model. We use the model to investigate two questions:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Does the direction of the diversity-stability relationship remain
  positive when species coexistence is fluctuation-dependent? And,
\item
  When species coexistence is fluctuation-dependent, how does increasing
  environmental variability impact ecosystem stability?
\end{enumerate}

@@

Eisenhauer, N., A. D. Barnes, S. Cesarz, D. Craven, O. Ferlian, F.
Gottschall, J. Hines, A. Sendek, J. Siebert, M. P. Thakur, and M.
T{ü}rke. 2016. Biodiversity-ecosystem function experiments reveal the
mechanisms underlying the consequences of biodiversity change in real
world ecosystems. Journal of Vegetation Science 27:1061--1070.

@@

@@

Wardle, D. A. 2016. Do experiments exploring plant diversity-ecosystem
functioning relationships inform how biodiversity loss impacts natural
ecosystems? Journal of Vegetation Science 27:646--653.
